### 2021-07-28 doc/thesis/chapters/hls.tex
sample rate reduction hls

     ...
 }
\end{CppCode}
\end{program}

>>>>

     auto input = sample_in.read();
     ...
     audio_out.write(audio);
 }
\end{CppCode}
\end{program}

HLS stream interface variables can be accessed via read- and write-functions.
One very important property of these functions is, that they are blocking calls.
The HLS streams are implemented with an underlying FIFO.
Thus, values can only be written into it when there is an empty spot, and can only be read if there is a at least one value available.
If the respective condition is not satisfied, the functions block the execution, until the FIFO state changes.
During simulation in the testbench, this may result in an endlessly blocking behaviour.

### 2021-07-28 doc/thesis/chapters/implementation.tex
sample rate reduction hls

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Digital Signal Processing Chain}

@@

The design uses an AXI4-Lite memory mapped interface, as well as multiple AXI4-Streams.
Additionally, a data-only interface is used for the connected LEDs.
The actual implementation is demonstrated in the following lines of code.\\

\begin{CppCode}
void fm_receiver_hls(hls::stream<iq_sample_t>& iq_in,
                     hls::stream<audio_sample_t>& audio_out,
                     config_t& config,
                     status_t* status,
                     ap_int<NUM_LEDS>* led_out) {

@@

  ...
}
\end{CppCode}

These few lines of code already implement the entire functionality of each interface type.
The streams use the types \texttt{iq\_sample\_t} and \texttt{audio\_sample\_t}, which are structs of two 16-bit fixed-point variables and thus represent a width of 32 bit.
The configuration and status registers are also implemented as structs, that contain multiple variables to represent the separate registers.
For the registers, a special detail is the distinction between the register types read-write and read-only.
This difference is made by using a value-by-reference or a pointer value, for configuration and status interface, respectively.
Consequently, the configuration registers become read-writeable and the status registers remain read-only.
The \texttt{bundle} keyword in the pragma-directives combine the two register groups into a single AXI4-Lite bus interface, instead of creating one bus each.
This may be of advantage in the application software later, since it only needs to handle a single base address.
The interface for the LEDs that can be connected is implemented as a data-only type, which does not have any protocol assigned.
It simply represents the data value that is written to the \cplusplus\ variable.
The usage of a data type with variable bitwidth, such as \texttt{ap\_int} in the example, is of advantage since only the actually required number of connections is implemented in the interface.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Digital Signal Processing Chain}
\label{sec:impl:dsp-chain}

@@

The design uses an AXI4-Lite memory mapped interface, as well as multiple AXI4-Streams.
Additionally, a data-only interface is used for the connected LEDs.
The actual implementation is demonstrated in Program \ref{prog:hls_impl_interfaces}.
These few lines of code already implement the entire functionality of each interface type.\\

\begin{program}
  \caption{Implementation of sample rate reduction, using for-loops and AXI stream variables with their underlying FIFO behaviour.}
  \label{prog:hls_impl_interfaces}
\begin{CppCode}
void fm_receiver_hls(hls::stream<iq_sample_t>& iq_in,
                     hls::stream<audio_sample_t>& audio_out,
                     config_t& config,
                     status_t* status,
                     ap_int<NUM_LEDS>* led_out) {

@@

  ...
}
\end{CppCode}
\end{program}

The streams use the types \texttt{iq\_sample\_t} and \texttt{audio\_sample\_t}, which are structs of two 16-bit fixed-point variables and thus represent a width of 32 bit.
The configuration and status registers are also implemented as structs, that contain multiple variables to represent the separate registers.
For the registers, a special detail is the distinction between the register types read-write and read-only.
This difference is made by using a value-by-reference or a pointer value, for configuration and status interface, respectively.
Consequently, the configuration registers become read-writeable and the status registers remain read-only.\\

The \texttt{bundle} keyword in the pragma-directives combine the two register groups into a single AXI4-Lite bus interface, instead of creating one bus each.
This may be of advantage in the application software later, since it only needs to handle a single base address.
The interface for the LEDs that can be connected is implemented as a data-only type, which does not have any protocol assigned.
It simply represents the data value that is written to the \cplusplus\ variable.
The usage of a data type with variable bitwidth, such as \texttt{ap\_int} in the example, is of advantage since only the actually required number of connections is implemented in the interface.

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sample Rate Reduction}

Multiple steps in the DSP chain require the operation of sample rate reduction.
This is explained in section \ref{sec:impl:dsp-chain}.\\

In the VHDL implementation, this is implemented with a counter that counts the number of incoming samples.
If the counter reaches the number that matches the downsampling rate, is forwards the current sample.
Otherwise, the incoming samples are discarded.\\

In the HLS implementation however, this is done in another way.
Here the previous block, which still runs in the oversampled domain, is executed N times, where N is the oversampling rate, before a single output sample is forwarded.
This behaviour is possible because of the FIFO that is implemented in an AXI stream interface.\\

The implementation of the sample rate reduction can be shown in a short code example in Program \ref{prog:sample_rate_reduction_fifo_behaviour}.
The FM Demodulator runs in the oversampled domain and is executed \texttt{OSR\_RX} times in the inner for-loop.
Only the last sample is written to the \texttt{fm\_channel\_data} variable, which itself is an AXI stream type with an underlying FIFO depth of \texttt{OSR\_AUDIO}, as defined with the accompanying pragma.
Thus, the inner for-loop is executed \texttt{OSR\_AUDIO} times by the outer loop, to fill this FIFO until it is full.
The subsequent Channel Decoder then reads the entire FIFO and also implements downsampling with a similar behaviour.
Several samples are processed, where only the last one is forwarded to the subsequent processing step.
The FIFO behaviour of AXI stream variables is described in Section \ref{sec:hls:coding:interfaces}.

\begin{program}
  \caption{Implementation of sample rate reduction, using for-loops and AXI stream variables with their underlying FIFO behaviour.}
  \label{prog:sample_rate_reduction_fifo_behaviour}
\begin{CppCode}
audio_sample_t fm_receiver(hls::stream<iq_sample_t>& iq_in) {

  hls::stream<sample_t> fm_channel_data;
  #pragma HLS STREAM depth = OSR_AUDIO variable = fm_channel_data

  sample_t fm_demod;
  for (uint8_t i = 0; i < OSR_AUDIO; i++) {
    for (uint8_t k = 0; k < OSR_RX; k++) {
      iq_sample_t iq = iq_in.read();
      fm_demod       = fm_demodulator(iq);

      if (k == OSR_RX - 1) {
        fm_channel_data.write(fm_demod);
      }
    }
  }

  audio_sample_t audio_sample = channel_decoder(fm_channel_data);

  return audio_sample;
}

audio_sample_t channel_decoder(hls::stream<sample_t>& in_sample) {
  sample_t audio_mono;

  for (uint8_t i = 0; i < OSR_AUDIO; i++) {
    sample_t sample = in_sample.read();

    audio_mono = recover_mono(sample);
    ...
  }
  ...
  return audio;
}
\end{CppCode}
\end{program}
